Quality Criteria for SKOS Vocabularies

### 2012-03-20 criteria.tex
fixed intro to 3

We identified an initial set of quality criteria for SKOS vocabularies by reviewing existing literature and manually examining existing vocabularies. We published our findings online\footnote{\url{https://github.com/cmader/qSKOS/wiki/Quality-Criteria-for-SKOS-Vocabularies}} and requested feedback from experts via public mailing list and informal face to face discussions. Based on the responses we received, we then selected a subset of these criteria and formalized them into computable \textbf{quality checking functions}. Each function operates on a SKOS vocabulary, which we define, for the purpose of this work, as follows:

\begin{definition}[SKOS Vocabulary] Let the fully entailed RDFS interpretation~\cite{RDFSEM2012} (enriched by entailment of \texttt{owl:inverseOf} properties \todo{CM}{why?}) of a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with

\begin{itemize}
	\item \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of resources,

	\item $LV = I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of literal values,

	\item \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{conceptual resources},

	\item \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating conceptual resources with one another.

\end{itemize}

@@

The commonality of all quality checking functions is that they take a given SKOS vocabulary as input and return a (nested) set, containing the resources that do \todo{CM}{not - sure?} fulfill the quality criterion. Therefore, we can formalize a quality criterion as being an abstract quality checking function $f:V \rightarrow \mathcal{P}(IR)$, mapping a given vocabulary $v \in V$ to the powerset of its resources. Each quality criterion we found can be implemented as a realization of such an abstract filtering function and statistical analysis can be computed on top of the return values.

@@

>>>>

We identified an initial set of quality criteria for SKOS vocabularies by reviewing existing literature and manually examining existing vocabularies. We published our findings online\footnote{\url{https://github.com/cmader/qSKOS/wiki/Quality-Criteria-for-SKOS-Vocabularies}} and requested feedback from experts via public mailing list and informal face to face discussions. Based on the received responses, we then selected a subset of these criteria and formalized them into computable \textbf{quality checking functions}. Each function operates on a SKOS vocabulary, which we define, for the purpose of this work, as follows:

\begin{definition}[SKOS Vocabulary] Let a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of \textbf{resources}, \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{conceptual resources}, $LV = I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of \textbf{literal values}, and \item \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating conceptual resources with one another. Further, we let $V$ be the fully entailed RDFS interpretation~\cite{RDFSEM2012} of the underlying RDF graph, enriched by entailment of \texttt{owl:inverseOf} properties.

@@

The commonality of all quality checking functions is that they take a given SKOS vocabulary as input and return a (nested) set, containing the resources fulfill the quality criterion. Therefore, we can formalize a quality criterion as being an abstract quality checking function $f:V \rightarrow \mathcal{P}(IR)$, mapping a given vocabulary $v \in V$ to the powerset of its resources. Each quality criterion we found can be implemented as a realization of such an abstract filtering function and statistical analysis can be computed on top of the return values.

@@

The criteria defined in this section focus
